### 2022-09-30 Draft_Mid_Term.Rmd
Draft_Midterm_Rmarkdown

---
title: "Draft_Mid_Term"
author: "Yiping Ying"
date: "2022-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set Up:
```{r load_packages, include=FALSE}

# You can set some global options for knitting chunks

knitr::opts_chunk$set(echo = TRUE)

# Load some libraries

library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot) # plot correlation plot
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots

# functions and data directory

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#25CB10", "#5AB60C", "#8FA108",   "#C48C04", "#FA7800")
```

```{r set_directory,include=FALSE}
setwd("C:/Users/yyp/Desktop/Upenn/MUSA/courses/MUSA5080/Midterm")
```

```{r raw_data}
#原始数据
data<-st_read("./data/studentData.geojson")
```

```{r challenge_set}
#challenge_set:用于最后预测的组，crs=4326
challenge_set<-data %>% filter(price==0)
```

```{r model_set}
#model_set:建模组,crs=4326
model_set<-data %>% filter(price!=0)
```

```{r nhoods}
#读取该地区的地图,crs=4326(?ESRI:103500)
nhoods<-st_read('./data/neighborhood/npa.shp')

nhoods<-nhoods %>% st_transform(crs=4326)
```

```{r nhoods_plot}
#作图,坐标4326
ggplot()+
  geom_sf(data=nhoods)
```

```{r crime_data}
crime<-st_read("./data/Crime - Violent.geojson")
```

```{r crime_data_plot}
ggplot()+
  geom_sf(data=crime,aes(fill=X2011),color="transparent")
```