### 2022-10-03 Draft_Mid_Term.Rmd
10/03 update
knn added

---
title: "Draft_Mid_Term"
author: "Yiping Ying"
date: "2022-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# II.Data Manipulation and Visualization

## 2.0 Set Up
```{r load_packages, include=FALSE}

# You can set some global options for knitting chunks

knitr::opts_chunk$set(echo = TRUE)

# Load some libraries

library(tidyverse)
library(sf)
library(spdep)
library(caret)
library(ckanr)
library(FNN)
library(grid)
library(gridExtra)
library(ggcorrplot) # plot correlation plot
library(corrr)      # another way to plot correlation plot
library(kableExtra)
library(jtools)     # for regression model plots
library(ggstance) # to support jtools plots
library(ggpubr)    # plotting R^2 value on ggplot point scatter
library(broom.mixed) # needed for effects plots
library(stargazer)
library(tidycensus)

options(scipen = 999)
options(tigris_class = "sf")

# functions and data directory

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#0571b0","#92c5de","#ffffbf","#f4a582","#ca0020")
```

```{r set_directory,include=FALSE}
setwd("C:/Users/yyp/Desktop/Upenn/MUSA/courses/MUSA5080/Midterm")
```

```{r load_key, warning = FALSE,message=FALSE}
census_api_key("c1c749a5ad57fa58fe19e1403d894e9acef41a97", install=TRUE,overwrite = TRUE)
```
## 2.1 Data Wrangling

### 2.1.1 Data loading
```{r raw_data}
#Raw Data:
data<-st_read("./data/studentData.geojson")

#Add age,pricepersq,bathrooms
data<-data %>% mutate(AGE = 2022-yearbuilt,
                                pricepersq = price/shape_Area,
                                bathrooms = fullbaths+halfbaths)
```

```{r download_data,warning=FALSE,message=FALSE}
#download open data,crs=4326(?ESRI:103500)

#Census Tract:
tract<-st_read('./data/CensusTract/Census_Tracts_2020.shp')
tract<-tract %>% st_transform(crs=4326)

#Zipcode:
##用于neighborhood effects分析
zipcode<-st_read('./data/zipcode/Zipcode.shp')
zipcode<-zipcode %>% st_transform(crs=4326) %>% rename(shapeLeng = shape_Leng,shapeArea = shape_Area) %>% dplyr::select(zip,po_name,shapeLeng,shapeArea)
```

```{r}
ggplot()+
  geom_sf(data=st_union(zipcode))+
  geom_sf(data=park_amenities.sf)
```

```{r}
ggplot()+
  geom_sf(data=st_union(tract))+
  geom_sf(data=park_easement.sf)
```

```{r amenity_facility_feature,warning=FALSE}
#1.college:
college<-st_read('./data/college/Colleges.shp')
college.sf<-college %>% st_transform(crs = 4326) %>% dplyr::select(geometry) %>% na.omit() 
#2.daycare:
daycare<-st_read('./data/daycare/Daycare.shp')
daycare.sf<-daycare %>% st_transform(crs=4326)%>% dplyr::select(geometry) %>% na.omit()

#3.Medical Facilities:
MedicalFacilities<-st_read('./data/medical/MedicalFacilities.shp')
MedicalFacilities.sf<-MedicalFacilities %>% st_transform(crs=4326)%>% dplyr::select(geometry) %>% na.omit()

#4.library:
## 数目不太多
library<-st_read('./data/library/Library.shp')
library.sf<-library %>% st_transform(crs=4326)%>% dplyr::select(geometry) %>% na.omit()

#5.Historical cemeteries:
## 原始数据是polygon,为了方便计算，使用st_centroid取了每个polygon的质心
hist_cemeteries<-st_read('./data/historic_cemeteries/Historic_Cemeteries.shp')
hist_cemeteries.sf<-hist_cemeteries %>% st_transform(crs=4326)%>% st_centroid() %>% dplyr::select(geometry) %>% na.omit()

#6.Historical Properties:
## 原始数据是polygon,为了方便计算，使用st_centroid取了每个polygon的质心
hist_properties<-st_read('./data/historic_properties/HistoricProperty_National.shp')
hist_properties.sf<-hist_properties%>% st_transform(crs=4326)%>%st_centroid() %>% dplyr::select(geometry) %>% na.omit()

#7.church
## 数目非常多
church<-st_read('./data/church/Churches.shp')
church.sf<-church %>% st_transform(crs=4326)%>% dplyr::select(geometry) %>% na.omit()

#8.park_amenities:
park_amenities<-st_read('./data/park_amenities/ParkAmenity.shp')
park_amenities.sf<-park_amenities%>% st_transform(crs=4326)%>% dplyr::select(geometry) %>% na.omit()

#9.park easement:
park_easement<-st_read('./data/park_easement/ParkEasements.shp')
park_easement.sf<-park_easement%>% st_transform(crs=4326)%>%st_centroid() %>% dplyr::select(geometry) %>% na.omit()

#10.Capital Improvement Projects:
capital_improve<-st_read('./data/Capital_Improvement_Projects/Capital_Improvement_Projects.shp')
capital_improve.sf<-capital_improve %>% st_transform(crs=4326)%>% dplyr::select(geometry) %>% na.omit()

#11.Business Investment Opportunity Zones:
##这个=做binary指标。原始数据是polygon;考虑使用buffer?
business_invest<-st_read('./data/Business/Business_Investment_Opportunity_Zones.shp')
business_invest<-business_invest %>% st_transform(crs=4326)
```

```{r neighborhood_effects_factor_and_clean}
#spatial features collection
##这个chunk内的都是以census tract为单位的指标；将合并到data

#1.Crime:
crime<-st_read('./data/Crime - Violent.geojson')
crime<-crime %>% st_transform(crs=4326)
crime1<-crime %>% mutate(crime_sum = X2019.raw+X2020.raw+X2021.raw) %>%  dplyr::select(crime_sum)

#2.Long Commute:
long_commute<-st_read('./data/LongCommute.geojson')
long_commute<-long_commute %>% st_transform(crs=4326) %>% mutate(PctLongCommute = X2020) %>%  dplyr::select(PctLongCommute)

#3.Percent of Grocery:
grocery<-st_read('./data/Proximity to a Grocery Store.geojson')
Pctgrocery<-grocery %>% st_transform(crs=4326) %>% mutate(Pctgrocery = X2021) %>%  dplyr::select(Pctgrocery)

#4.Percent of canopy in residential area:
canopy<-st_read('./data/Tree Canopy - Residential.geojson')
Pctcanopy<-canopy %>% st_transform(crs=4326) %>% mutate(Pctcanopy = X2012) %>% dplyr::select(Pctcanopy)

#5.Percent of Test Proficiency in Elementary Schools:
test_elementary<-st_read('./data/Test Proficiency - Elementary School.geojson')
Pcttest1<-test_elementary%>% st_transform(crs=4326) %>%mutate( test.elementary = X2019) %>% dplyr::select(test.elementary)

#6.Percent of Test Proficiency in Middle Schools:
test_middle<-st_read('./data/Test Proficiency - Middle School.geojson')
Pcttest2<-test_middle%>% st_transform(crs=4326) %>%mutate( test.middle = X2019) %>% dplyr::select(test.middle)

#7.Percent of Test Proficiency in High Schools:
test_high<-st_read('./data/Test Proficiency - High School.geojson')
Pcttest3<-test_high%>% st_transform(crs=4326) %>% mutate(test.high = X2019) %>% dplyr::select(test.high)

#8.Percent of Bachelor's Degree:
bachelor<-st_read('./data/Education Level - Bachelor Degree.geojson')
PctBachelor<-bachelor %>% st_transform(crs=4326) %>% mutate(PctBachelor = X2020) %>% dplyr::select(PctBachelor)

#9.Median Age of Residents:
age<-st_read('./data/Age of Residents.geojson')
MedAge<-age %>% st_transform(crs=4326) %>% mutate(MedAge = X2020) %>% dplyr::select(MedAge)
```

Census Data:
```{r census_data,include=FALSE}
options(tigris_use_cache = TRUE)

acs_variable_list.2020 <- load_variables(2020, 
                                         "acs5", 
                                         cache = TRUE)

tracts20 <- 
  get_acs(geography = "tract", variables = c("B25026_001E","B02001_002E", "B19013_001E","B25058_001E","B06012_002E",                            
"B25002_003E","B25004_002E","B25004_003E","B25004_004E","B25004_005E",
"B25001_001E","B08137_003E"), 
          year=2020, state= 37, county= 119, geometry=T, output="wide")%>% st_transform(crs=4326) %>%
  rename(TotalPop = B25026_001E, 
         Whites = B02001_002E,
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E,
         TotalPoverty = B06012_002E,
         TotalVacant = B25002_003E,
         ForRent = B25004_002E,
         ForRentVac = B25004_003E,
         ForSale = B25004_004E,
         ForSaleVac = B25004_005E,
         TotalUnit = B25001_001E,
         RenterOccupied = B08137_003E
         ) %>%
  dplyr::select(-NAME, -starts_with("B")) %>% #-starts_with("B") awesome!
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop,0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         pctTotalVacant = ifelse(TotalUnit > 0, TotalVacant / TotalUnit * 100, 0),
         TotalOccupied = TotalUnit - TotalVacant,
         pctRenterOccupied = ifelse(TotalOccupied >0, RenterOccupied/TotalOccupied, 0)) %>%
  dplyr::select(-Whites, -TotalPoverty) 
```

### 2.1.2 Feature Engineering

```{r}
ggplot()+
  geom_sf(data=zipcode)
```

```{r data_left_join}
data<-st_join(data,crime1)
data<-st_join(data,long_commute)
data<-st_join(data,Pctgrocery)
data<-st_join(data,Pctcanopy)
data<-st_join(data,Pcttest1)
data<-st_join(data,Pcttest2)
data<-st_join(data,Pcttest3)
data<-st_join(data,PctBachelor)
data<-st_join(data,MedAge)
data<-st_join(data,zipcode)
data<-st_join(data,tracts20)
```

```{r knn}
st_c<-st_coordinates

#1.college:
data <-
  data %>% 
  mutate(
    college_nn1 = nn_function(st_c(data),st_c(college.sf), 1),
    college_nn2 = nn_function(st_c(data),st_c(college.sf), 2),
    college_nn3 = nn_function(st_c(data),st_c(college.sf), 3),
    college_nn4 = nn_function(st_c(data),st_c(college.sf), 4),
    college_nn5 = nn_function(st_c(data),st_c(college.sf), 5),
    )

#2.daycare:
data <-
  data %>% 
  mutate(
    daycare_nn1 = nn_function(st_c(data),st_c(daycare.sf), 1),
    daycare_nn2 = nn_function(st_c(data),st_c(daycare.sf), 2),
    daycare_nn3 = nn_function(st_c(data),st_c(daycare.sf), 3),
    daycare_nn4 = nn_function(st_c(data),st_c(daycare.sf), 4),
    daycare_nn5 = nn_function(st_c(data),st_c(daycare.sf), 5),
    )

#3.Medical Facilities:
data <-
  data %>% 
  mutate(
    MedicalFacilities_nn1 = nn_function(st_c(data),st_c(MedicalFacilities.sf), 1),
    MedicalFacilities_nn2 = nn_function(st_c(data),st_c(MedicalFacilities.sf), 2),
    MedicalFacilities_nn3 = nn_function(st_c(data),st_c(MedicalFacilities.sf), 3),
    MedicalFacilities_nn4 = nn_function(st_c(data),st_c(MedicalFacilities.sf), 4),
    MedicalFacilities_nn5 = nn_function(st_c(data),st_c(MedicalFacilities.sf), 5),
    )

#4.library:
data <-
  data %>% 
  mutate(
    library_nn1 = nn_function(st_c(data),st_c(library.sf), 1),
    library_nn2 = nn_function(st_c(data),st_c(library.sf), 2),
    library_nn3 = nn_function(st_c(data),st_c(library.sf), 3),
    library_nn4 = nn_function(st_c(data),st_c(library.sf), 4),
    library_nn5 = nn_function(st_c(data),st_c(library.sf), 5),
    )

#5.Historical cemeteries:
data <-
  data %>% 
  mutate(
    hist_cemeteries_nn1 = nn_function(st_c(data),st_c(hist_cemeteries.sf), 1),
    hist_cemeteries_nn2 = nn_function(st_c(data),st_c(hist_cemeteries.sf), 2),
    hist_cemeteries_nn3 = nn_function(st_c(data),st_c(hist_cemeteries.sf), 3),
    hist_cemeteries_nn4 = nn_function(st_c(data),st_c(hist_cemeteries.sf), 4),
    hist_cemeteries_nn5 = nn_function(st_c(data),st_c(hist_cemeteries.sf), 5),
    )

#6.Historical Properties:
data <-
  data %>% 
  mutate(
    hist_properties_nn1 = nn_function(st_c(data),st_c(hist_properties.sf), 1),
    hist_properties_nn2 = nn_function(st_c(data),st_c(hist_properties.sf), 2),
    hist_properties_nn3 = nn_function(st_c(data),st_c(hist_properties.sf), 3),
    hist_properties_nn4 = nn_function(st_c(data),st_c(hist_properties.sf), 4),
    hist_properties_nn5 = nn_function(st_c(data),st_c(hist_properties.sf), 5),
    )

#7.church
data <-
  data %>% 
  mutate(
    church_nn1 = nn_function(st_c(data),st_c(church.sf), 1),
    church_nn2 = nn_function(st_c(data),st_c(church.sf), 2),
    church_nn3 = nn_function(st_c(data),st_c(church.sf), 3),
    church_nn4 = nn_function(st_c(data),st_c(church.sf), 4),
    church_nn5 = nn_function(st_c(data),st_c(church.sf), 5),
    )

#9.park easement:
data <-
  data %>% 
  mutate(
    park_easement_nn1 = nn_function(st_c(data),st_c(park_easement.sf), 1),
    park_easement_nn2 = nn_function(st_c(data),st_c(park_easement.sf), 2),
    park_easement_nn3 = nn_function(st_c(data),st_c(park_easement.sf), 3),
    park_easement_nn4 = nn_function(st_c(data),st_c(park_easement.sf), 4),
    park_easement_nn5 = nn_function(st_c(data),st_c(park_easement.sf), 5),
    )

#10.Capital Improvement Projects:
data <-
  data %>% 
  mutate(
    capital_improve_nn1 = nn_function(st_c(data),st_c(capital_improve.sf), 1),
    capital_improve_nn2 = nn_function(st_c(data),st_c(capital_improve.sf), 2),
    capital_improve_nn3 = nn_function(st_c(data),st_c(capital_improve.sf), 3),
    capital_improve_nn4 = nn_function(st_c(data),st_c(capital_improve.sf), 4),
    capital_improve_nn5 = nn_function(st_c(data),st_c(capital_improve.sf), 5),
    )
```

```{r eval=FALSE}
#8.park_amenities:
###这个需要将三围点变成二维点
data0 <-
  data0 %>% 
  mutate(
    park_amenities_nn1 = nn_function(st_c(data0),st_c(park_amenities.sf), 1),
    park_amenities_nn2 = nn_function(st_c(data0),st_c(park_amenities.sf), 2),
    park_amenities_nn3 = nn_function(st_c(data0),st_c(park_amenities.sf), 3),
    park_amenities_nn4 = nn_function(st_c(data0),st_c(park_amenities.sf), 4),
    park_amenities_nn5 = nn_function(st_c(data0),st_c(park_amenities.sf), 5),
    )
```

### 2.1.3 Split dataset
```{r challenge_set}
#challenge_set:用于最后预测的组，crs=4326
challenge_set<-data %>% filter(price==0)
```

```{r model_set}
#model_set:建模组,crs=4326
model_set<-data %>% filter(price!=0)
```

## 2.2 Summary Statistics of Features

model_set internal features
```{r summary_internal_features}
#Summary and selection of internal features:
internal.features<-model_set%>% dplyr::select(price,shape_Area,pricepersq,AGE,heatedarea,numfirepla,fullbaths,halfbaths,bedrooms,units) %>% st_drop_geometry()

#stargazer:
stargazer(internal.features, type = "html", 
          title = "Table DATA 2.1 Summary Statistics of Internal Characteristics ",
          header = FALSE,
          single.row = TRUE)
```

```{r internal_feature_correlation_matrix}
#Correlation matrix of internal features:
numericVars <- 
  select_if(internal.features, is.numeric) %>% na.omit()

ggcorrplot(
  round(cor(numericVars), 1), 
  p.mat = cor_pmat(numericVars),
  colors = c("#0571b0", "#ffffff", "#ca0020"),
  type="lower",
  insig = "pch") +  
    labs(title = "Correlation across numeric variables")+plotTheme()
```
```{r summary_amenities_feature}
amenities_feature.list<-data %>% dplyr::select(starts_with('daycare_'),starts_with('MedicalFacilities_'),starts_with('hist_c'),starts_with('park_'),starts_with('capital_'),starts_with('Pct'),starts_with('test.e'),pctTotalVacant, pctRenterOccupied, pctWhite,pctPoverty,MedRent ,MedHHInc,MedAge) %>% st_drop_geometry()

#stargazer:
stargazer(amenities_feature.list, type = "html", 
          title = "Table DATA 2.1 Summary Statistics of Internal Characteristics ",
          header = FALSE,
          single.row = TRUE)
```

```{r amenities_feature_matrix}
numericVars2 <- 
  select_if(amenities_feature.list, is.numeric) %>% na.omit()

ggcorrplot(
  round(cor(numericVars2), 1), 
  p.mat = cor_pmat(numericVars2),
  colors = c("#0571b0", "#ffffff", "#ca0020"),
  type="lower",
  insig = "pch") +  
    labs(title = "Correlation across numeric variables")+plotTheme()
```

## Map Home Price
```{r plot_home_price}
#Plot of home price per square feet:
ggplot()+
  geom_sf(data=zipcode,fill="grey40",color="white")+
  geom_sf(data=model_set,aes(color=q5(pricepersq)),show.legend="point",size=.75)+
  scale_color_manual(values = palette5,labels = qBr(model_set, "pricepersq"),name = "House Price\nPer Square Feet\n(Quintile Breaks)")+
  labs(title="House Price",subtitle = "Mecklenberg County,NC",caption="Figure1.1")+
  mapTheme()
```